---
title: "BYSY project Notebook"
output: html_notebook
---  
# 03_RFD Profiles  

***  
## Computing forks coverages and RFD

```{r}
suppressMessages(library(rtracklayer))
suppressMessages(library(tidyverse))
suppressMessages(library(GenomicRanges))
`%+%` <- paste0
seqinfSY <- readRDS("Data/seqinfSY.rds")
seqinfBY <- readRDS("Data/seqinfBY.rds")
```

There are two pairs of diverging Ty element from the chrIII and chrXVI that generate basecalling/mapping artifacts. We will therefore remove the forks, initiations and terminations mapping on these genomic segments.  
```{r}
Ty_chrIII_SY <- GRanges("CP029160.1",IRanges(7409106,7421212),strand="*",seqinfo=seqinfSY)
Ty_chrIII_BY <- GRanges("CP026297.1",IRanges(174645,186751),strand="*",seqinfo=seqinfBY)
Ty_chrXVI_SY <- GRanges("CP029160.1",IRanges(851297,863439),strand="*",seqinfo=seqinfSY)
Ty_chrXVI_BY <- GRanges("CP026290.1",IRanges(844705,856847),strand="*",seqinfo=seqinfBY)
maskedTy_SY <- c(Ty_chrIII_SY,Ty_chrXVI_SY)
saveRDS(maskedTy_SY, file="Data/maskedTy_SY.rds")
maskedTy_BY <- c(Ty_chrIII_BY,Ty_chrXVI_BY)
saveRDS(maskedTy_BY, file="Data/maskedTy_BY.rds")
```

### Importing forks and removing those mapping on the problematic Ty area.
```{r}
forks_SYSY <- readRDS("Data/forks_SYSY.rds")
forks_SYSY.gr<- with(forks_SYSY, GRanges(seqnames=chrom,ranges=IRanges(pmin(X0,X1),pmax(X0,X1)),strand=case_when(direc=="R"~"+",T~"-"),seqinfo=seqinfSY))
forks_SYSY2 <- forks_SYSY %>% 
	mutate(Ty_filter=!overlapsAny(forks_SYSY.gr,maskedTy_SY)) %>%
	filter(Ty_filter)
forks_SYSY2.gr<- with(forks_SYSY2, GRanges(seqnames=chrom,ranges=IRanges(pmin(X0,X1),pmax(X0,X1)),strand=case_when(direc=="R"~"+",T~"-"),seqinfo=seqinfSY))

forks_BYSY <- readRDS("Data/forks_BYBYSY.rds")
forks_BYSY.gr<- with(forks_BYSY, GRanges(seqnames=chromSY,ranges=IRanges(pmin(X0nSY,X1nSY),pmax(X0nSY,X1nSY)),strand=case_when(direc=="R"~"+",T~"-"),seqinfo=seqinfSY))
forks_BYSY2 <- forks_BYSY %>% 
	mutate(Ty_filter=!overlapsAny(forks_BYSY.gr,maskedTy_SY)) %>%
	filter(Ty_filter)
forks_BYSY2.gr<- with(forks_BYSY2, GRanges(seqnames=chromSY,ranges=IRanges(pmin(X0nSY,X1nSY),pmax(X0nSY,X1nSY)),strand=case_when(direc=="R"~"+",T~"-"),seqinfo=seqinfSY))

forks_BYBY <- readRDS("Data/forks_BYBY.rds")
forks_BYBY.gr<- with(forks_BYBY, GRanges(seqnames=chrom,ranges=IRanges(pmin(X0,X1),pmax(X0,X1)),strand=case_when(direc=="R"~"+",T~"-"),seqinfo=seqinfBY))
forks_BYBY2 <- forks_BYBY %>% 
	mutate(Ty_filter=!overlapsAny(forks_BYBY.gr,maskedTy_BY)) %>%
	filter(Ty_filter)
forks_BYBY2.gr<- with(forks_BYBY2, GRanges(seqnames=chrom,ranges=IRanges(pmin(X0,X1),pmax(X0,X1)),strand=case_when(direc=="R"~"+",T~"-"),seqinfo=seqinfBY))
```

We can then compute the RFD for SY data mapped on SY genome,BY data mapped on BY genome and transposed to SY genome and BY data mapped on BY genome.  
```{r}
## on SY genome
forks_SYSY_R.gr<- forks_SYSY2.gr[strand(forks_SYSY2.gr)=="+"]
forks_SYSY_L.gr<- forks_SYSY2.gr[strand(forks_SYSY2.gr)=="-"]

covL_SY <- coverage(forks_SYSY_L.gr)
covR_SY <- coverage(forks_SYSY_R.gr)
cov_SY <- covR_SY+covL_SY
rfdSY <- (covR_SY-covL_SY)/(covR_SY+covL_SY)
export(covL_SY,con="BigWig/covL_SYSY_nt.bw")
export(covR_SY,con="BigWig/covR_SYSY_nt.bw")
export(cov_SY,con="BigWig/covtot_SYSY_nt.bw")
export(rfdSY,con="BigWig/rfd_SYSY_nt.bw")

# BY mapped on BY transposed to SY
forks_BYSY_R.gr<- forks_BYSY2.gr[strand(forks_BYSY2.gr)=="+"]
forks_BYSY_L.gr<- forks_BYSY2.gr[strand(forks_BYSY2.gr)=="-"]

covL_BY <- coverage(forks_BYSY_L.gr)
covR_BY <- coverage(forks_BYSY_R.gr)
cov_BY <- covR_BY+covL_BY
rfdBY <- (covR_BY-covL_BY)/(covR_BY+covL_BY)
export(covL_BY,con="BigWig/covL_BYBYSY_nt.bw")
export(covR_BY,con="BigWig/covR_BYBYSY_nt.bw")
export(cov_BY,con="BigWig/covtot_BYBYSY_nt.bw")
export(rfdBY,con="BigWig/rfd_BYBYSY_nt.bw")

## on BY genome
forks_BYBY_R.gr<- forks_BYBY2.gr[strand(forks_BYBY2.gr)=="+"]
forks_BYBY_L.gr<- forks_BYBY2.gr[strand(forks_BYBY2.gr)=="-"]

covL_BY <- coverage(forks_BYBY_L.gr)
covR_BY <- coverage(forks_BYBY_R.gr)
rfdBY <- (covR_BY-covL_BY)/(covR_BY+covL_BY)
export(rfdBY,con="BigWig/rfd_BYBY_nt.bw")
```
