---
title: "BYSY project Notebook"
output: html_notebook
---  
# 11_Speed Map  

***  

To evaluate the differences in replication fork speed between the SY and BY strain, we proceed similarly as in [Theulot et al, 2022](https://doi.org/10.1038/s41467-022-31012-0).  
Forks were reduced to their center. We choose to remove forks mapping on the  problematic Ty loci from chrIII and chrXVI and the ura3∆0 locus. 
The SY genome was parted into non overlapping 1kb bins on which forks were affected with respect to their center.  
When we computed the forks speed distribution for genomic bins above 1kb, we choose to proceed by merging consecutive 1kb bins, keeping a 1kb shift.  
As an illustration, the bin 73 at the 3kb scale contains the forks mapping on the bins 72,73 and 74 from the 1kb scale.
For all this procedure, data were pooled by strain.  

```{r message=FALSE, warning=FALSE}
suppressMessages(library(GenomicRanges))
suppressMessages(library(tidyverse))
suppressMessages(library(rtracklayer))
library(future.apply)
library(parallel)
library(furrr)
library(patchwork)
library(ggprism)
library(DescTools)
`%+%`<-paste0
seqinfSY <- readRDS("Data/seqinfSY.rds")
rDNA_SY <- GRanges("CP029160.1",IRanges(3879940,3934000),strand="*",seqinfo=seqinfSY)
maskedTy_SY <- readRDS("Data/maskedTy_SY.rds")
masked_ura3d0 <- GRanges("CP029160.1",IRanges(1051200,1051500),strand="*",seqinfo=seqinfSY,type="ura3d0")
masked_SY <- c(maskedTy_SY,masked_ura3d0) 

source("Helper_function.r")
theme_set(theme_bw())
mypal <- c(paletteer::paletteer_d("ggthemes::Classic_20"),"grey40","black")
```

### Importing forks
```{r message=FALSE, warning=FALSE, cache=TRUE}
forksBY <- readRDS("Data/forks_BYBYSY.rds")
forksSY <- readRDS("Data/forks_SYSY.rds")

fBY <- forksBY %>%
	select(chrom=chromSY,direc,X0=X0nSY,X1=X1nSY,exp,speed,type) %>%
	mutate(Rep=map_chr(exp,function(x) x %>% str_remove("BY_"))) %>%
	mutate(strain="BY")
fSY <- forksSY %>%
	select(chrom,direc,X0,X1,exp,speed,type) %>%
	mutate(Rep=map_chr(exp,function(x) x %>% str_remove("SY_"))) %>%
	mutate(strain="SY")


forkBY.gr<- with(fBY, GRanges(seqnames=chrom,ranges=IRanges(pmin(X0,X1),pmax(X0,X1)),strand=case_when(direc=="L"~"+",T~"-"),seqinfo=seqinfSY,speed=speed,direc=direc,type=type,Rep=Rep))
## exclude TychrIII, TychrXVI and ura3d0 forks
forkBY.gr <- forkBY.gr[!overlapsAny(forkBY.gr,masked_SY)]
speedmedgenBY <- median(forkBY.gr$speed,na.rm=T)
# 2402
forkSY.gr<- with(fSY, GRanges(seqnames=chrom,ranges=IRanges(pmin(X0,X1),pmax(X0,X1)),strand=case_when(direc=="L"~"+",T~"-"),seqinfo=seqinfSY,speed=speed,direc=direc,type=type,Rep=Rep))
forkSY.gr <- forkSY.gr[!overlapsAny(forkSY.gr,masked_SY)]
speedmedgenSY <- median(forkSY.gr$speed,na.rm=T)
# 2432
forkBY.grc <- resize(forkBY.gr,fix="center",width=1)
forkSY.grc <- resize(forkSY.gr,fix="center",width=1)
```

### Affecting speeds to genomic bins

```{r message=FALSE, warning=FALSE, cache=TRUE}
bs <- 1000 
bingen <- tileGenome(seqinfSY,tilewidth=bs, cut.last.tile.in.chrom=T)
nc <- 8L

ol1 <- findOverlaps(forkBY.grc,bingen)
bingen$BY1k <- mclapply(1:length(bingen), function(x) tibble(speed=forkBY.grc[queryHits(ol1)[subjectHits(ol1)==x]]$speed,direc=forkBY.grc[queryHits(ol1)[subjectHits(ol1)==x]]$direc,type=forkBY.grc[queryHits(ol1)[subjectHits(ol1)==x]]$type,Rep=forkBY.grc[queryHits(ol1)[subjectHits(ol1)==x]]$Rep),mc.cores=nc)

ol1 <- findOverlaps(forkSY.grc,bingen)
bingen$SY1k <- mclapply(1:length(bingen), function(x) tibble(speed=forkSY.grc[queryHits(ol1)[subjectHits(ol1)==x]]$speed,direc=forkSY.grc[queryHits(ol1)[subjectHits(ol1)==x]]$direc,type=forkSY.grc[queryHits(ol1)[subjectHits(ol1)==x]]$type,Rep=forkSY.grc[queryHits(ol1)[subjectHits(ol1)==x]]$Rep),mc.cores=nc)
```

### Generating other scales binning
```{r message=FALSE, warning=FALSE, cache=TRUE}
roll_c_mc <- function(x,win=3,nc=4L)
{
# x is a list
	require(future.apply)
	plan(multicore, workers = nc)

	res0 <- future_lapply(1:(length(x)-win+1), function(i) {
		purrr::reduce(x[i:(i+win-1)], dplyr::full_join, by = c("speed","direc","type","Rep"))
		})
	if (win%%2==0) 
	{res <- c(rep(list(NA),win%/%2-1),res0,rep(list(NA),win%/%2))
	}else{res <- c(rep(list(NA),win%/%2),res0,rep(list(NA),win%/%2))
	}
return(res)
}

bingen$BY3k <- roll_c_mc(bingen$BY1k,win=3)
bingen$SY3k <- roll_c_mc(bingen$SY1k,win=3)
#saveRDS(bingen,"centerfullpooledSpeed_" %+% data_suff %+% ".rds")
bingen$BY5k <- roll_c_mc(bingen$BY1k,win=5)
bingen$SY5k <- roll_c_mc(bingen$SY1k,win=5)
#saveRDS(bingen,"centerfullpooledSpeed_" %+% data_suff %+% ".rds")
bingen$BY11k <- roll_c_mc(bingen$BY1k,win=11)
bingen$SY11k <- roll_c_mc(bingen$SY1k,win=11)
#saveRDS(bingen,"centerfullpooledSpeed_" %+% data_suff %+% ".rds")
bingen$BY21k <- roll_c_mc(bingen$BY1k,win=21)
bingen$SY21k <- roll_c_mc(bingen$SY1k,win=21)
#saveRDS(bingen,"centerfullpooledSpeed_" %+% data_suff %+% ".rds")
bingen$BY41k <- roll_c_mc(bingen$BY1k,win=41)
bingen$SY41k <- roll_c_mc(bingen$SY1k,win=41)
saveRDS(bingen,"Data/centerfullpooledSpeed.rds")
```

### Exporting 21k speed and coverage

```{r message=FALSE, warning=FALSE, cache=TRUE}
bingen <- readRDS("Data/centerfullpooledSpeed.rds")
speedBY21k <- lapply(bingen$BY21k, function(x) {
	if (!is.na(x)[1])
	{res <- summarize(x,speedmed=MedianCI(speed)[1],speedmedCIinf=MedianCI(speed)[2],speedmedCIsup=MedianCI(speed)[3],n=n())} else {res <- tibble(speedmed=NA,speedmedCIinf=NA,speedmedCIsup=NA,n=NA)}
	return(res)
	}) 
speedBY21k2 <- do.call(bind_rows,speedBY21k)
speedSY21k <- lapply(bingen$SY21k, function(x) {
	if (!is.na(x)[1])
	{res <- summarize(x,speedmed=MedianCI(speed)[1],speedmedCIinf=MedianCI(speed)[2],speedmedCIsup=MedianCI(speed)[3],n=n())} else {res <- tibble(speedmed=NA,speedmedCIinf=NA,speedmedCIsup=NA,n=NA)}
	return(res)
	}) 
speedSY21k2 <- do.call(bind_rows,speedSY21k)

export(coverage(bingen,weight=speedBY21k2$n),con="BigWig/BYforkcovC21k.bw")
export(coverage(bingen,weight=speedBY21k2$speedmed),con="BigWig/BYforkspeedmedC21k.bw")
export(coverage(bingen,weight=speedSY21k2$n),con="BigWig/SYforkcovC21k.bw")
export(coverage(bingen,weight=speedSY21k2$speedmed),con="BigWig/SYforkspeedmedC21k.bw")
```

### By strain comparison

We choose to test for each strain and for each bin if the speed distribution in this bin is different from the whole genome speed distribution. rDNA speeds were excluded from the whole genome distribution.
```{r message=FALSE, warning=FALSE, cache=TRUE}
	require(future.apply)
	plan(multicore, workers = 10L)

bingenNoR <- bingen[!overlapsAny(bingen,rDNA_SY)]
# BY
speed.binBY <- do.call(c,lapply(bingenNoR$BY1k,function(x) x%>%pull(speed)))
BYtest <- bingen
BYtest$wtl1k <- future_sapply(BYtest$BY1k, function(x) {
		test <- nrow(x)!=0
		if (length(test)==0) {test=F}
		if (test) {		
			res <- wilcox.test(x$speed,speed.binBY,alternative="l")$p.value} else {res<- NA}
	return(res)
	})
BYtest$wtg1k <- 1-BYtest$wtl1k
BYtest$wtl3k <- future_sapply(BYtest$BY3k, function(x) {
		test <- nrow(x)!=0
		if (length(test)==0) {test=F}
		if (test) {		
			res <- wilcox.test(x$speed,speed.binBY,alternative="l")$p.value} else {res<- NA}
	return(res)
	})
BYtest$wtg3k <- 1-BYtest$wtl3k
BYtest$wtl5k <- future_sapply(BYtest$BY5k, function(x) {
		test <- nrow(x)!=0
		if (length(test)==0) {test=F}
		if (test) {		
			res <- wilcox.test(x$speed,speed.binBY,alternative="l")$p.value} else {res<- NA}
	return(res)
	})
BYtest$wtg5k <- 1-BYtest$wtl5k
BYtest$wtl11k <- future_sapply(BYtest$BY11k, function(x) {
		test <- nrow(x)!=0
		if (length(test)==0) {test=F}
		if (test) {		
			res <- wilcox.test(x$speed,speed.binBY,alternative="l")$p.value} else {res<- NA}
	return(res)
	})
BYtest$wtg11k <- 1-BYtest$wtl11k
BYtest$wtl21k <- future_sapply(BYtest$BY21k, function(x) {
		test <- nrow(x)!=0
		if (length(test)==0) {test=F}
		if (test) {		
			res <- wilcox.test(x$speed,speed.binBY,alternative="l")$p.value} else {res<- NA}
	return(res)
	})
BYtest$wtg21k <- 1-BYtest$wtl21k
BYtest$wtl41k <- future_sapply(BYtest$BY41k, function(x) {
		test <- nrow(x)!=0
		if (length(test)==0) {test=F}
		if (test) {		
			res <- wilcox.test(x$speed,speed.binBY,alternative="l")$p.value} else {res<- NA}
	return(res)
	})
BYtest$wtg41k <- 1-BYtest$wtl41k
saveRDS(BYtest,file="Data/BYspeedtest_centered.rds")
# SY
speed.binSY <- do.call(c,lapply(bingenNoR$SY1k,function(x) x%>%pull(speed)))

SYtest <- bingen
SYtest$wtl1k <- future_sapply(SYtest$SY1k, function(x) {
		test <- nrow(x)!=0
		if (length(test)==0) {test=F}
		if (test) {		
			res <- wilcox.test(x$speed,speed.binSY,alternative="l")$p.value} else {res<- NA}
	return(res)
	})
SYtest$wtg1k <- 1-SYtest$wtl1k
SYtest$wtl3k <- future_sapply(SYtest$SY3k, function(x) {
		test <- nrow(x)!=0
		if (length(test)==0) {test=F}
		if (test) {		
			res <- wilcox.test(x$speed,speed.binSY,alternative="l")$p.value} else {res<- NA}
	return(res)
	})
SYtest$wtg3k <- 1-SYtest$wtl3k
SYtest$wtl5k <- future_sapply(SYtest$SY5k, function(x) {
		test <- nrow(x)!=0
		if (length(test)==0) {test=F}
		if (test) {		
			res <- wilcox.test(x$speed,speed.binSY,alternative="l")$p.value} else {res<- NA}
	return(res)
	})
SYtest$wtg5k <- 1-SYtest$wtl5k
SYtest$wtl11k <- future_sapply(SYtest$SY11k, function(x) {
		test <- nrow(x)!=0
		if (length(test)==0) {test=F}
		if (test) {		
			res <- wilcox.test(x$speed,speed.binSY,alternative="l")$p.value} else {res<- NA}
	return(res)
	})
SYtest$wtg11k <- 1-SYtest$wtl11k
SYtest$wtl21k <- future_sapply(SYtest$SY21k, function(x) {
		test <- nrow(x)!=0
		if (length(test)==0) {test=F}
		if (test) {		
			res <- wilcox.test(x$speed,speed.binSY,alternative="l")$p.value} else {res<- NA}
	return(res)
	})
SYtest$wtg21k <- 1-SYtest$wtl21k
SYtest$wtl41k <- future_sapply(SYtest$SY41k, function(x) {
		test <- nrow(x)!=0
		if (length(test)==0) {test=F}
		if (test) {		
			res <- wilcox.test(x$speed,speed.binSY,alternative="l")$p.value} else {res<- NA}
	return(res)
	})
SYtest$wtg41k <- 1-SYtest$wtl41k
saveRDS(SYtest,file="Data/SYspeedtest_centered.rds")

### holm correction for the pvalues
BYwtl_adj <- as_tibble(BYtest) %>% 
	select(seqnames,start,end,strand,contains("wtl")) %>% 
	pivot_longer(contains("wtl")) %>%
	mutate(padj=p.adjust(value,"holm")) %>% 
	select(-"value") %>%
	pivot_wider(names_from = name,values_from=c(padj),names_prefix="adj_")
BYwtg_adj <- as_tibble(BYtest) %>% 
	select(seqnames,start,end,strand,contains("wtg")) %>% 
	pivot_longer(contains("wtg")) %>%
	mutate(padj=p.adjust(value,"holm")) %>% 
	select(-"value") %>%
	pivot_wider(names_from = name,values_from=c(padj),names_prefix="adj_")

BYwt_adj <- left_join(BYwtl_adj,BYwtg_adj)
saveRDS(BYwt_adj,file="Data/BYspeedtestadjHolm_centered.rds")

SYwtl_adj <- as_tibble(SYtest) %>% 
	select(seqnames,start,end,strand,contains("wtl")) %>% 
	pivot_longer(contains("wtl")) %>%
	mutate(padj=p.adjust(value,"holm")) %>% 
	select(-"value") %>%
	pivot_wider(names_from = name,values_from=c(padj),names_prefix="adj_")
SYwtg_adj <- as_tibble(SYtest) %>% 
	select(seqnames,start,end,strand,contains("wtg")) %>% 
	pivot_longer(contains("wtg")) %>%
	mutate(padj=p.adjust(value,"holm")) %>% 
	select(-"value") %>%
	pivot_wider(names_from = name,values_from=c(padj),names_prefix="adj_")

SYwt_adj <- left_join(SYwtl_adj,SYwtg_adj)
saveRDS(SYwt_adj,file="Data/SYspeedtestadjHolm_centered.rds")

```

#### SY vs BY stat test

```{r message=FALSE, warning=FALSE, cache=TRUE}
bg <- as_tibble(bingen)

bg2 <- bg %>%
	mutate(wtl1k=pmap_dbl(.,function(SY1k,BY1k,...) {
		if (nrow(SY1k)!=0 & nrow(BY1k)!=0) {
			x <- SY1k %>% pull(speed)
			y <- BY1k %>% pull(speed)
			res <- wilcox.test(x,y,alternative="l")$p.value
			}else{
			res<- NA}
		return(res)
		})) %>%
	mutate(wtg1k=1-wtl1k)
	
bg3 <- bg2 %>%
	mutate(wtl3k=future_pmap_dbl(.,function(SY3k,BY3k,...) {
		test <- (nrow(SY3k)!=0 & nrow(BY3k)!=0)
		if (length(test)==0) {test=F}
		if (test) {
			x <- SY3k %>% pull(speed)
			y <- BY3k %>% pull(speed)
			res <- wilcox.test(x,y,alternative="l")$p.value
			}else{
			res<- NA}
		return(res)
		})) %>%
	mutate(wtg3k=1-wtl3k)%>%
	mutate(wtl5k=future_pmap_dbl(.,function(SY5k,BY5k,...) {
		test <- (nrow(SY5k)!=0 & nrow(BY5k)!=0)
		if (length(test)==0) {test=F}
		if (test) {
			x <- SY5k %>% pull(speed)
			y <- BY5k %>% pull(speed)
			res <- wilcox.test(x,y,alternative="l")$p.value
			}else{
			res<- NA}
		return(res)
		})) %>%
	mutate(wtg5k=1-wtl5k)%>%
	mutate(wtl11k=future_pmap_dbl(.,function(SY11k,BY11k,...) {
		test <- (nrow(SY11k)!=0 & nrow(BY11k)!=0)
		if (length(test)==0) {test=F}
		if (test) {
			x <- SY11k %>% pull(speed)
			y <- BY11k %>% pull(speed)
			res <- wilcox.test(x,y,alternative="l")$p.value
			}else{
			res<- NA}
		return(res)
		})) %>%
	mutate(wtg11k=1-wtl11k)%>%
	mutate(wtl21k=future_pmap_dbl(.,function(SY21k,BY21k,...) {
		test <- (nrow(SY21k)!=0 & nrow(BY21k)!=0)
		if (length(test)==0) {test=F}
		if (test) {
			x <- SY21k %>% pull(speed)
			y <- BY21k %>% pull(speed)
			res <- wilcox.test(x,y,alternative="l")$p.value
			}else{
			res<- NA}
		return(res)
		})) %>%
	mutate(wtg21k=1-wtl21k)%>%
	mutate(wtl41k=future_pmap_dbl(.,function(SY41k,BY41k,...) {
		test <- (nrow(SY41k)!=0 & nrow(BY41k)!=0)
		if (length(test)==0) {test=F}
		if (test) {
			x <- SY41k %>% pull(speed)
			y <- BY41k %>% pull(speed)
			res <- wilcox.test(x,y,alternative="l")$p.value
			}else{
			res<- NA}
		return(res)
		})) %>%
	mutate(wtg41k=1-wtl41k)
	
saveRDS(bg3,file="Data/centerfullpooledSpeed_Wtest.rds")


### Holm correction for the pvalues
wtl_adj <- bg3 %>% 
	select(seqnames,start,end,strand,contains("wtl")) %>% 
	pivot_longer(contains("wtl")) %>%
	mutate(padj=p.adjust(value,"holm")) %>% 
	select(-"value") %>%
	pivot_wider(names_from = name,values_from=c(padj),names_prefix="adj_")
wtg_adj <- bg3 %>% 
	select(seqnames,start,end,strand,contains("wtg")) %>% 
	pivot_longer(contains("wtg")) %>%
	mutate(padj=p.adjust(value,"holm")) %>% 
	select(-"value") %>%
	pivot_wider(names_from = name,values_from=c(padj),names_prefix="adj_")

res_tot <- left_join(bg3,wtl_adj) %>% left_join(wtg_adj)

saveRDS(res_tot,file="Data/centerfullpooledSpeed_Wtest_Holmpadj.rds")
```

